\ttitle

### 2013-05-24 Appendices/AppendixA-summaries.tex
Finish SVM summary
Summary CUSUM method

\subsubsection{SVCPD: The algorithm}
Instead of using statistical properties of the data, for each window of size $w$ a hyper-sphere is constructed without increasing computational complexity due to the \emph{kernel trick}.
The window size is related to the sensitivity of the method to change; small windows are sensitive with high false alarm rate whilst large windows are slow to detect change and have low alarm rates.
The algorithm is:

\begin{center}\begin{table}
\begin{tabular}{ l p{12cm} }
  \hline
  Step & Action \\
  \hline
  1 & Start with $n$ observations and construct hyper-sphere \\
  2 & Add next observation $x_t$ and drop first one \\
  3 & Identify new hyper-sphere and its \emph{approximate radius} \\
  4 & if $x_t$ is outside hyper-sphere, mark $t$ as change points and continue from step 2 \\
  5 & calculate radius average of last $w$ hyper-planes \\
  6 & calculate radius ratio $\hbar$. If lower than ${th}_{low}$ or greater than ${th}_{high}$ then mark $t$ as change point \\
  7 & continue from step 2 \\
  \hline
\end{tabular}
\caption{Support Vector machine based Change Point Detection algorithm}
\end{table}\end{center}

>>>>

\subsubsection{SVCPD: The algorithm}
Instead of using statistical properties of the data, for each window of size $w$ a hyper-sphere is constructed without increasing computational complexity due to the \emph{kernel trick}.
The window size is related to the sensitivity of the method to change; small windows are sensitive with high false alarm rate whilst large windows are slow to detect change and have low alarm rates.
The algorithm is listed in table \ref{tab:alg-svcdp}.
Note that SVCPD can be applied directly to multidimensional data, whilst many other methods can only be applied to one-dimensional data.

\begin{center}\begin{table}
\begin{tabular}{ l p{12cm} }
  \hline
  Step & Action \\
  \hline
  1 & Start with $n$ observations and construct hyper-sphere \\
  2 & Add next observation $x_t$ and drop first one \\
  3 & Identify new hyper-sphere and its \emph{approximate radius} \\
  4 & if $x_t$ is outside hyper-sphere, mark $t$ as change points and continue from step 2 \\
  5 & calculate radius average of last $w$ hyper-planes \\
  6 & calculate radius ratio $\hbar$. If lower than ${th}_{low}$ or greater than ${th}_{high}$ then mark $t$ as change point \\
  7 & continue from step 2 \\
  \hline
\end{tabular}
\caption{Support Vector machine based Change Point Detection algorithm}
\label{tab:alg-svcdp}
\end{table}\end{center}

%----------------------------------------------------------------------------------------
\section{CUSUM for variance}

\subsection{Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance}
Carla Inclan and George C. Tiao \cite{inclan1994use}, 1944, 162 refs.

%--------------------------------------------
\subsubsection{Introduction}
This paper is about reflective detection of multiple changes of variance in a sequence of independent observations.
This is a statistical method, which differs from others (in that field) such as Bayesian method (Bayes ratio, posterior odds), maximum likelihood methods and (autoregressive) models.
This approach uses the centered version of cumulative sums of squares to search for change points systematically and iteratively (and reflective).

%--------------------------------------------
\subsubsection{Centered Cumulative Sum of Squares}
The cumulative sum of squares is often used for change detection in the mean.
It is defined as $C_k = \sum_{i=1}^k \alpha_t^2$ for a series of uncorrelated random variables $\{\alpha_t\}$ with mean $0$ and variance $\sigma_t^2, t = 1, 2, \dots, T$.
The centered (and normalized) cumulative sum of squares is:
\begin{equation}
  \begin{aligned}
  D_k = \frac{C_k}{C_T} - \frac{k}{T}, & & k = 1, \dots, T, & & \text{with } D_0 = D_T = 0
  \end{aligned}
\end{equation}
For homogeneous variance the plot of $D_k$ against $k$ (the first $k$ elements of the series) will oscillate around $0$.
When a sudden change in variance occurs, the pattern of the plot of $D_k$ will break out some specified boundaries with high probability.
For $C_k$ it holds that, under homogeneous variance, the plot will be a straight line with slope $\sigma^2$.
With one of more change points the plot appears as a line of several straight pieces.
The plot of $D_k$ creates a peak for a smaller and a trough for a larger variance, is visually more clear and breaks out a predefined value.
The search for a change point is variance is than to find $k^* = \text{max}_k \abs{D_k}$.
If the value of $D_k$ at $k^*$ exceeds a predefined value (e.g. $D_{0.5}^* = 1.358$, for $\sqrt{T / 2D_k}$ because of the Brownian bridge property), that value of $k^*$ will be an estimate for a change point.

There is a relation between $D_k$ and the $F$ statistic, which is used for testing equality of variances between two independent samples.
For a fixed $k$, $D_k(F)$ is a monotone function of $F$ (it depends only on $k$ through $k/T$).
An important distinction: the $F$ statistic is \emph{used with known $k$}, whereas we are looking for $\text{max}_k \abs{D_k}$ to determine the location of the change point.

When we assume that $\{\alpha_t\}$ is normally distributed with mean $0$ and variances $\sigma_t^2$, then we can obtain the \emph{likelihood ratio} for testing the hypothesis of one change against the hypothesis of no change in the variance.
When maximizing the likelihood estimator for a location $\kappa$, we can find the log-likelihood ratio $LR_{0,1}$.
Although $LR_{0,1}$ and $\text{max}_k \abs{D_k}$ are related, they are not the same.
The latter puts more weight near the middle of the series is thus biased toward $T/2$.

The (expected) value of $D_k$ given a change in variance differs in the context.
If a smaller variance corresponds to the smaller portion of the series, then it will be harden to find the change point using $D_k$.
There is a masking effect when there are multiple change points in the series; the order of small, medium and large variances result in the value of $D_k$.
The iterative algorithm presented in this paper in section \ref{subsec:icss} is designed to lessen the masking effect.

%--------------------------------------------
\subsubsection{Multiple changes: Iterated cumulative sums of squares}\label{subsec:icss}
In case of a single change point the $D_k$ method would succeed.
But we are interested in multiple change points of variance, and thus the usefulness of the $D_k$ reduces due to the masking effect.
A solution is to iteratively applying the method and dividing the series at each possible change point.
The algorithm is presented in table \ref{tab:icss}.
It is the third steps which reduces the masking effect and helps to ``fine tune'' the algorithm by (re)moving the potential change points by checking each location given the adjacent ones.

\begin{center}\begin{table}
\begin{tabular}{ l p{12cm} }
  \hline
  Step & Action \\
  \hline
  0 & Let $t_1 = 1$ \\
  1 & Calculate $D_k(\alpha[t_1 : T])$. Let $k^*(\alpha[t_1 : T])$ be the point at which $\text{max}_k \abs{D_k(\alpha[t_1 : T])}$. Let $D^*$ be the asymptotically critical value and $M$ the max value in the series segment (?). If $M > D^*$ then consider $k^*$ to be a change point. Else, there is no change point and the algorithm stops. \\
  2a & Repeat for the first part (up to the change point), until no more change points are found. \\
  2b & Repeat for the second part (from the change point forward), until no more change points are found. \\
  3 & When two or more change points are found; check for each $\alpha[j-1 : j+1]$ if there is indeed a change point ($j$). Repeat until the number of change points does not change and each new found change point is ``close'' enough to previous. \\
  \hline
\end{tabular}
\caption{Iterated Cumulative Sums of Squares (ICSS) Algorithm}
\label{tab:alg-svcdp}
\end{table}\end{center}

\subsubsection{Results}
When the ICSS algorithm was applied to stock data, it resulted in comparable results as the maximum likelihood estimates and Bayesian analysis.
The performance (CPU-time and correct observations with artificial data) of ICSS outperforms the other two.
The heavy computational burden of posterior odds can be partially alleviated by the maximum log-likelihood method.
The ICSS algorithm avoids calculating a function at all possible locations of change points due the iterative manner.
